---
title: "Power Analysis Jan 18"
author: "Danielle Galvin"
date: "1/18/2022"
output: html_document
---

```{r}
library(tidyverse)
library(brms)
```

Data simulation:
NEED TO UPDATE THE MEANS AND SDS SO THAT THEY ARE REFLECTIVE OF THE LITERATURE
Based on the 2010 paper by Swansons and Burdick on cricket frog overwintering, the mean values were as follows with the sample size in parentheses:

Glucose is in units of micromoles / gram wet mass
Glycogen is in units of micromole glucosyl units / gram wet mass

liver glucose: control (2.5 +/- 1.6 (5)), 6-hour-frozen (16.50 +/- 3.55 (8)), 24-hour-frozen (8.8 +/- 5.4 (8))

liver glycogen: control (157.1 +/- 48.8 (6)), 6-hour-frozen (256.0 +/- 131.1 (8)), 24-hour-frozen (113.9 +/- 159.6 (8))

muscle glucose: control (1.0 +/- 0.5 (6)), 6-hour-frozen (0.79 +/- 0.48 (8)), 24-hour-frozen (1.3 +/- 1.2 (8))

muscle glycogen: control (6.9 +/- 5.3 (6)), 6-hour-frozen (16.81 +/- 9.7 (8)), 24-hour-frozen (16.8 +/- 22.0 (8))
```{r}
#I am choosing to use the values from the control groups for my data analysis since I do not anticipate them being frozen. This should provide a good baseline as I am just looking to determine if the total glucose is different based on the different treatments. Treatment 1 represents the normal mesocosms while treatment 2 will represent the insulated mesocosms.

#Treatment 1 simulations for liver glucose:
#three_trt1 <- tibble(glucose=rnorm(3, 2.5, 1.6)) %>% 
#  mutate(trt="trt1")
#four_trt1 <- tibble(glucose=rnorm(4, 2.5, 1.6)) %>% 
#  mutate(trt="trt1")
five_trt1 <- tibble(glucose=rnorm(5, 2.5, 1.6)) %>% 
  mutate(trt="trt1")
six_trt1 <- tibble(glucose=rnorm(6, 2.5, 1.6)) %>% 
  mutate(trt="trt1")
seven_trt1 <- tibble(glucose=rnorm(7, 2.5, 1.6)) %>% 
  mutate(trt="trt1")
#eight_trt1 <- tibble(glucose=rnorm(8, 2.5, 1.6)) %>% 
#  mutate(trt="trt1")

#Treatment 2 simulations for liver glucose where I am going to make the values lower because the insulated mesocosms should be warmer on average:
#three_trt2 <- tibble(glucose=rnorm(3, 1.5, 2)) %>% 
 # mutate(trt="trt2")
#four_trt2 <- tibble(glucose=rnorm(4, 1.5, 2)) %>% 
 # mutate(trt="trt2")
five_trt2 <- tibble(glucose=rnorm(5, 1.5, 2)) %>% #Try running with a lower standard deviation to imply a strong effect whereas currently I am implying a weak effect
  mutate(trt="trt2")
six_trt2 <- tibble(glucose=rnorm(6, 1.5, 2)) %>% 
  mutate(trt="trt2")
seven_trt2 <- tibble(glucose=rnorm(7, 1.5, 2)) %>% 
  mutate(trt="trt2")
eight_trt2 <- tibble(glucose=rnorm(8, 1.5, 2)) %>% 
  mutate(trt="trt2")


#Class code:
test <- tibble(trt1=rnorm(8, 2.5, 2),
               trt2=rnorm(8, 1.5, 0.2)) %>% 
  pivot_longer(cols=everything())

test %>% 
  ggplot(aes(x=name, y=value))+
  geom_boxplot(aes(group=name))+
  geom_point()+
  ylim(-5,5)



#Making the combinations I wish to evaluate
#three_three <- bind_rows(three_trt1, three_trt2)
#three_five <- bind_rows(three_trt1, five_trt2)
#five_three <- bind_rows(five_trt1, three_trt2)
five_five <- bind_rows(five_trt1, five_trt2)
#five_seven <- bind_rows(five_trt1, seven_trt2)
#seven_five <- bind_rows(seven_trt1, five_trt2)
#six_eight <- bind_rows(six_trt1, eight_trt2)
#eight_six <- bind_rows(eight_trt1, six_trt2)
six_six <- bind_rows(six_trt1, six_trt2)
#six_seven <- bind_rows(six_trt1, seven_trt2)
#seven_six <- bind_rows(seven_trt1, six_trt2)
seven_eight <- bind_rows(seven_trt1, eight_trt2)
seven_seven <- bind_rows(seven_trt1, seven_trt2)


```

Sensitivity analysis:
```{r}
seven_eight_brm <- brm(glucose ~ trt,
                 data=seven_eight,
                 family=gaussian(),
                 file="Models/seven_eight_brm.rds",
                 file_refit="on_change")
                
saveRDS(seven_eight_brm, "Models/seven_eight_brm.rds")

#eight_seven_brm <- brm(glucose ~ 1,
 #                data = eight_seven,
  #               family = gaussian(),
   #              file="Models/eight_seven_brm.rds",
    #             file_refit="on_change")

#saveRDS(eight_seven_brm, "Models/eight_seven_brm.rds")

seven_seven_brm <- brm(glucose ~ trt,
                       data = seven_seven,
                       family=gaussian(),
                       file="Models/seven_seven_brm.rds",
                       file_refit="on_change")

saveRDS(seven_seven_brm, "Models/seven_seven_brm.rds")

#three_three_brm <- brm(glucose ~ 1,
                     #  data=three_three,
                      # family=gaussian(),
                       #file="Models/three_three_brm.rds",
                       #file_refit="on_change")

#saveRDS(three_three_brm, "Models/three_three_brm.rds")

five_five_brm <- brm(glucose ~ trt,
                     data=five_five,
                     family=gaussian(),
                     file="Models/five_five_brm.rds",
                     file_refit="on_change")

saveRDS(five_five_brm, "Models/five_five_brm.rds")

six_six_brm <- brm(glucose ~ trt,
                   data=six_six,
                   family=gaussian(),
                   file="Models/six_six_brm.rds",
                   file_refit="on_change")

saveRDS(six_six_brm, "Models/six_six_brm.rds")

#six_seven_brm <- brm(glucose ~ 1,
 #                    data=six_seven,
  #                   family=gaussian(),
   #                  file="Models/six_seven_brm.rds",
    #                 file_refit="on_change")

#saveRDS(six_seven_brm, "Models/six_seven_brm.rds")

#seven_six_brm <- brm(glucose ~ 1,
 #                    data=seven_six,
  #                   family=gaussian(),
   #                  file="Models/seven_six_brm.rds",
    #                 file_refit="on_change")

#saveRDS(seven_six_brm, "Models/seven_six_brm.rds")
```

```{r}
plot(seven_eight_brm)
#plot(eight_seven_brm)
plot(seven_seven_brm)
#plot(three_three_brm)
plot(five_five_brm)
plot(six_six_brm)
#plot(six_seven_brm)
#plot(seven_six_brm)
```






