---
title: "Analysis_Bayes_Class"
author: "Danielle Galvin"
output: html_document
---

WELCOME TO THE FUN ZONE

Loading our packages
```{r}
library(readr)
library(brms)
library(tidyverse)
library(rethinking)
library(janitor)
library(scales)
library(readxl)
library(tidybayes)
```

Bringing in the data and cleaning it up
```{r}
Data_full <- read_csv("Data/Data_full.csv") 

#PLEASE IGNORE, THERE IS NO WIZARD HIDING BEHIND THE CURTAIN

#I want to make the data long so that all measurements, stages, and lengths are in single categories and are associated with their dates. This chunk will make measure instance and date of measurement.
#data_long <- Data_full %>% pivot_longer(cols=-c(individual, treatment_group, species_num, species, stage_1, mass_1, length_1, stage_2, mass_2, length_2, stage_3, mass_3, length_3, stage_4, mass_4, length_4, stage_5, mass_5, length_5, `6/15/2021`, `6/16/2021`, `6/17/2021`, `6/18/2021`, `6/19/2021`, `6/20/2021`, `6/21/2021`, `6/22/2021`, `6/23/2021`, `6/24/2021`, `6/25/2021`, `6/26/2021`, `6/27/2021`, `6/28/2021`, `6/29/2021`, `6/30/2021`, `7/1/2021`, `7/2/2021`, `7/3/2021`, `7/4/2021`, `7/5/2021`, `7/6/2021`, `7/7/2021`, `7/8/2021`, `7/9/2021`, `7/10/2021`, `7/11/2021`, `7/12/2021`, `7/13/2021`, `7/14/2021`, `7/15/2021`, `7/16/2021`, `7/17/2021`, `7/18/2021`, `7/19/2021`, `7/20/2021`, `7/21/2021`, `7/22/2021`, `7/23/2021`, `7/24/2021`, `7/25/2021`, `7/26/2021`, `7/27/2021`, `7/28/2021`, `7/29/2021`, `7/30/2021`, `7/31/2021`, `8/1/2021`, `8/2/2021`, `8/3/2021`, `8/4/2021`, `8/5/2021`, `8/6/2021`, `8/7/2021`, `8/8/2021`, `8/9/2021`, `8/10/2021`, `8/11/2021`, `8/12/2021`, `8/13/2021`, `8/14/2021`, `8/15/2021`, `8/16/2021`, `8/17/2021`, `8/18/2021`, `8/19/2021`, `8/20/2021`, `8/21/2021`, `8/22/2021`, `8/23/2021`, `8/24/2021`, `8/25/2021`, `8/26/2021`, `8/27/2021`, `8/28/2021`, `8/29/2021`, `8/30/2021`, `8/31/2021`, `9/1/2021`, `9/2/2021`, `9/3/2021`, `9/4/2021`, `9/5/2021`, `9/6/2021`, `9/7/2021`, `9/8/2021`, `9/9/2021`, `9/10/2021`, `9/11/2021`, `9/12/2021`, `9/13/2021`), names_to="measure_instance", values_to="measure_date")

#Now I want to combine all alive dates and alive status to cut down on colums
#data_longer <- data_long %>% pivot_longer(cols=-c(individual, treatment_group, species_num, species, stage_1, mass_1, length_1, stage_2, mass_2, length_2, stage_3, mass_3, length_3, stage_4, mass_4, length_4, stage_5, mass_5, length_5, measure_instance, measure_date), names_to="alive_date", values_to="alive_status")

#Now I want to collapse the mass measures
#data_longer_2 <- data_longer %>% pivot_longer(cols=-c(individual, treatment_group, species_num, species, stage_1, length_1, stage_2, length_2, stage_3, length_3, stage_4, length_4, stage_5, length_5, measure_instance, measure_date, alive_date, alive_status), names_to="mass_instance", values_to="mass")

#Now I want to collapse stage measures
#data_longer_3 <- data_longer_2 %>% pivot_longer(cols=-c(individual, treatment_group, species_num, species, length_1, length_2, length_3, length_4, length_5, measure_instance, measure_date, alive_date, alive_status, mass_instance, mass), names_to="stage_instance", values_to="stage")

#Now collapse length
#data_longest <- data_longer_3 %>% pivot_longer(cols=-c(individual, treatment_group, species_num, species, measure_instance, measure_date, alive_date, alive_status, mass_instance, mass, stage_instance, stage), names_to="length_instance", values_to="length")

#Wait I think I messed up PLEASE ONLY LOOK AT THIS
data_guess <- Data_full %>% pivot_longer(cols=-c(individual, treatment_group, species_num, species, stage_1, mass_1, length_1, stage_2, mass_2, length_2, stage_3, mass_3, length_3, stage_4, mass_4, length_4, stage_5, mass_5, length_5, m_date1, m_date2, m_date3, m_date4, m_date5), names_to="date_status", values_to="alive_status")

#Adding a column for total days each individual survived
alive_data <- data_guess %>% group_by(individual) %>%  mutate(days_survived = sum(alive_status))

#Cutting the data frame down so my computer doesn't crash and ruin my life
alive_data_useme <- alive_data[, c("individual", "species_num", 
                                   "species", "days_survived", "treatment_group")] 

#Making control equal to 0
alive_data_useme$treatment_group[alive_data_useme$treatment_group == "control"] <-  0

#Okay really use this one!
real_data_alive <- alive_data_useme[!duplicated(alive_data_useme$individual),] %>% 
  transform(treatment_group=as.numeric(treatment_group)) %>% 
  mutate(trt_cat = as.character(treatment_group))
```

Model to investigate mortality date:
  y ~ dpois(lambda)
  log(lambda)  = a + bx
  a ~ ?
  b ~ ?

Simulate the priors
I am not currently happy with this. Based on the literature infection and mortality peaks at 10 days post exposure. How can I change the priors to reflect this information? I would expect that the peak would occur ealier based on a higher treatment group due to higher viral load.

REALLY NEED HELP WITH THIS AS MY PREDICTIONS ARE CURRENTLY GARBAGE
```{r}
N=1000 #Number of simulations

prior_sim <- tibble(a = rnorm(N, 0.05, 0.05),
                 b = rnorm(N, 0.05, 0.05),
                 sigma = rexp(N, 0.5),
                 sim = 1:N) 

# data (only the x values, since we're simulating y and mu and pretending we don't have them yet)
treatment_dg <- real_data_alive$treatment_group
d <- real_data_alive %>% mutate("treatment_sim" = treatment_dg)

# combine and simulate
prior_and_x_1 <- prior_sim %>% expand_grid(x_sim = treatment_dg) %>% # combine priors and x's
  mutate(lambda = exp(a + b*x_sim),                         # simulate regressions
         y = rpois(nrow(.), lambda=lambda))   # simulate data 
         
         
prior_and_x_1 %>% 
  ggplot(aes(x = x_sim, y = lambda, group = sim)) + 
  geom_line() +
  geom_point(aes(y = lambda)) +
  labs(y = "sim")+
  scale_y_log10()
```

Model to investigate mortality date
```{r}
#I am using a Poisson distribution since dates are integer values
d <- d %>% mutate("days" = days_survived)

get_prior(days ~ trt_cat*species, 
          data = d,
          family = poisson(link="log"))

bsr_brm_dg_prior <- brm(days ~ trt_cat*species, 
               data = d,
               family = poisson(link="log"),
               prior = c(prior(normal(-1, 0.5), class = "Intercept"),
                         prior(normal(-.75, 0.2), class = "b", coef="speciesr_pipiens"),
                         prior(normal(-0.75, 0.2), class = "b", coef="trt_cat10"),
                         prior(normal(-0.75, 0.2), class = "b", coef="trt_cat10:speciesr_pipiens"),
                         prior(normal(-0.75, 0.2), class = "b", coef="trt_cat100"),
                         prior(normal(-0.75, 0.2), class = "b", coef="trt_cat100:speciesr_pipiens"),
                         prior(normal(-0.75, 0.2), class = "b", coef="trt_cat1000"),
                         prior(normal(-0.75, 0.2), class = "b", coef="trt_cat1000:speciesr_pipiens"),
                         prior(exponential(0.5), class="b")), 
               cores = 1, chains = 1, iter = 1000,
               sample_prior = "only")

bsr_brm_dg_prior
conditional_effects(bsr_brm_dg_prior)


posteriors_bsr_brm_dg <- as_draws_df(bsr_brm_dg) %>% as_tibble() %>% 
  mutate(lambda = exp(b_Intercept),
         lambda_prior = exp(prior_Intercept))

ggplot(posteriors_bsr_brm_dg %>% filter(.draw <=1000), aes(x=b_treatment_sim, y=lambda_prior)) + 
  geom_point()
```

If you happen to be able to make the above data worth anything, I would appreciate if you had the opportunity to incorporate species into the model as well. Species number 1 is Pseudacris maculata (boreal chorus frog) which is very small and has no published data regarding survivability toward ranavirus. Species number 2 is Rana pipiens (Northern leopard frog) which is larger and has some published data regarding survivability to ranavirus, although results show that survival is not only dose dependent, but it is also dependent on the virulence of the specific strain. The strain I used was isolated from a Wood frog (Rana sylvatica) in Maine and the virulence in unknown to me.




To save a model, use the following code:
saveRDS(objectname, "Models/objectname.rds")

then, in your model code, after the iter line, add the following:
file="Models/objectname.rds",
file_refit="on_change")

This way the model will always be here and can more easily be updated without completely refitting.

```{r}
#running a predictive model in brms
d_new <- real_data_alive %>%
  mutate(trt_cat = as.character(treatment_group))

get_prior(days_survived ~ 1 + trt_cat*species,
family = Gamma(link="log"), data = d_new)

prior_test <- brm(days_survived ~ 1 + trt_cat*species,
                  family = Gamma(link="log"),
                  data = d_new,
                  prior = c(prior(normal(-0.3,0.1), class = "b", coef = speciesr_pipiens),
                            prior(normal(-0.2,0.1), class = "b"),
                            prior(normal(2.5,0.75), class = "Intercept")),
                  sample_prior = "only",
                  iter = 1000, chains = 1,
                  file="Models/prior_test.rds",
                  file_refit="on_change")

saveRDS(prior_test, "Models/prior_test.rds")

prior_test
plot(conditional_effects(prior_test)) # T shows your data points in the graph
```

```{r}
pp_check(prior_test)
pp_check(prior_test, type="stat")
pp_check(prior_test, type="stat_grouped", group="trt_cat")
pp_check(prior_test, type="stat_grouped", group="species")
```

The above model does not seem to be producing data that looks like mine, so I am making a new model with different values. Let's see if this looks better:
```{r}
#running a predictive model in brms
d_new <- real_data_alive %>%
  mutate(trt_cat = as.character(treatment_group))

get_prior(days_survived ~ 1 + trt_cat*species,
family = poisson(link="log"), data = d_new)

prior_take2 <- brm(days_survived ~ 1 + trt_cat*species,
                  family = poisson(link="log"),
                  data = d_new,
                  prior = c(prior(normal(-0.1,0.03), class = "b", coef = speciesr_pipiens),
                            prior(normal(-0.5,0.2), class = "b"),
                            prior(normal(2.8,0.5), class = "Intercept")),
                  #sample_prior = "only",
                  iter = 1000, chains = 1,
                  file="Models/prior_take2.rds",
                  file_refit="on_change")

saveRDS(prior_take2, "Models/prior_take2.rds")

prior_take2
plot(conditional_effects(prior_take2)) # T shows your data points in the graph

pp_check(prior_take2)
pp_check(prior_take2, type="stat")
pp_check(prior_take2, type="stat_grouped", group="trt_cat")
pp_check(prior_take2, type="stat_grouped", group="species")
```

That one was better but not quite it. Will continue to edit
```{r}
#running a predictive model in brms
d_new <- real_data_alive %>%
  mutate(trt_cat = as.character(treatment_group))

get_prior(days_survived ~ 1 + trt_cat*species,
family = poisson(link="log"), data = d_new)

prior_take3 <- brm(days_survived ~ 1 + trt_cat*species,
                  family = poisson(link="log"),
                  data = d_new,
                  prior = c(prior(normal(-0.001,0.0003), class = "b", coef = speciesr_pipiens),
                            prior(normal(-0.5,0.2), class = "b"),
                            prior(normal(2.8,0.5), class = "Intercept")),
                  #sample_prior = "only",
                  iter = 1000, chains = 1,
                  file="Models/prior_take3.rds",
                  file_refit="on_change")

saveRDS(prior_take3, "Models/prior_take3.rds")

prior_take3
plot(conditional_effects(prior_take3)) # T shows your data points in the graph

pp_check(prior_take3)
pp_check(prior_take3, type="stat")
pp_check(prior_take3, type="stat_grouped", group="trt_cat")
pp_check(prior_take3, type="stat_grouped", group="species")
```

I like this one, but by using default priors I can tell that the poisson likelihood is not the proper likelihood to use
```{r}
#running a predictive model in brms
d_new <- real_data_alive %>%
  mutate(trt_cat = as.character(treatment_group))

get_prior(days_survived ~ 1 + trt_cat*species,
family = poisson(link="log"), data = d_new)

prior_take4 <- brm(days_survived ~ 1 + trt_cat*species,
                  family = poisson(link="log"),
                  data = d_new,
                  #sample_prior = "only",
                  iter = 1000, chains = 1,
                  file="Models/prior_take4.rds",
                  file_refit="on_change")

prior_take4 <- update(prior_take4, chains=4)

saveRDS(prior_take4, "Models/prior_take4.rds")

prior_take4
plot(conditional_effects(prior_take4), points=T) # T shows your data points in the graph

pp_check(prior_take4)
pp_check(prior_take4, type="stat")
pp_check(prior_take4, type="stat_grouped", group="trt_cat")
pp_check(prior_take4, type="stat_grouped", group="species")


```

Code from Feb 1 class
```{r}
bayes_R2(prior_take4)
```

Try with a negative binomial dist instead of poisson
```{r}
#running a predictive model in brms
d_new <- real_data_alive %>%
  mutate(trt_cat = as.character(treatment_group))

get_prior(days_survived ~ 1 + trt_cat*species,
family = negbinomial(link="log", link_shape="log"), data = d_new)

final_model<- brm(days_survived ~ 1 + trt_cat*species,
                  family = negbinomial(link="log", link_shape="log"),
                  data = d_new,
                  prior = c(prior(normal(1,1.5), class = "b", coef = speciesr_pipiens),
                            prior(normal(-0.5,1.5), class = "b"),
                            prior(normal(2.8,3), class = "Intercept")),
                  #sample_prior = "only",
                  iter = 1000, chains = 4,
                  file="Models/final_model.rds",
                  file_refit="on_change")


saveRDS(final_model, "Models/final_model.rds")

final_model
plot(conditional_effects(final_model), points = T) # T shows your data points in the graph

pp_check(final_model)
pp_check(final_model, type="stat")
pp_check(final_model, type="stat_grouped", group="trt_cat")
pp_check(final_model, type="stat_grouped", group="species")
pp_check(final_model, type = "boxplot")
pp_check(final_model, type="dens_overlay_grouped", group="trt_cat")
pp_check(final_model, type="dens_overlay_grouped", group="species")


bayes_R2(final_model)


cond_data = distinct(final_model$data %>% select(-days_survived))

posts <- add_epred_draws(final_model, newdata = cond_data)

#This is the plot I want to use to show the difference in days survived by treatment group and species. The density is from the posterior and the dots are the raw data.
final_plot <- posts %>% 
  ggplot(aes(x = trt_cat, y = .epred, color = species)) + 
  geom_violin() +
  geom_point(data = final_model$data, aes(y = days_survived),
             position = position_dodge(width = 0.9)) +            #Position_dodge moves the dots to overlap the violins for a better visualization
  #scale_y_log10()+
  labs(y="Days Survived", x="Ranavirus Exposure (PFU/mL)")
final_plot
ggsave("bayesrvspeciestrtnonlog.jpg", plot=final_plot, width=6, height=4, units="in", dpi=600)


#Manipulating my data so that I can calculate the differences in treatment groups
pivoted_posts <- posts %>% 
  select(trt_cat, species, .draw,.row, .epred) %>% 
  ungroup() %>% 
  select(trt_cat, species, .draw, .epred) %>% 
  pivot_wider(id_cols=c(species,.draw),
              names_from=trt_cat,
              values_from=c( .epred)) %>% 
  pivot_wider(id_cols=.draw,
              names_from=species,
              values_from=c('0', '10', '100', '1000')) %>% 
  view()


#Calculating the differences
differences <- pivoted_posts %>% 
  mutate(diff_control=`0_r_pipiens`-`0_p_maculata`) %>%
  mutate(diff_tens=`10_r_pipiens` - `10_p_maculata`) %>% 
  mutate(diff_hundreds=`100_r_pipiens` - `100_p_maculata`) %>% 
  mutate(diff_thousands=`1000_r_pipiens` - `1000_p_maculata`)
  
#Calculating the median days survived and the 95% credible interval. There is a 95% probability that the values range from blank to blank with a median of blank.
differences %>% 
  median_qi(diff_thousands)
differences %>% 
  median_qi(diff_control)
differences %>% 
  median_qi(diff_tens)
differences %>% 
  median_qi(diff_hundreds)
  


#this calculates the probability that one group has a higher value than the other. In this case calculating the probability that r_pipiens survived more days than p_maculata. You can never have a straight 0 or 1 but instead could say there is less than a 0.01 probability or a greater than 0.99 probability.
differences %>% 
  summarize(higher = sum(diff_control>0)/nrow(.))
differences %>% 
  summarize(higher= sum(diff_tens>0)/nrow(.))
differences %>% 
  summarize(higher=sum(diff_hundreds>0)/nrow(.))
differences %>% 
  summarize(higher=sum(diff_thousands>0)/nrow(.))
  



ggplot(data=differences, aes(x=diff_control))+
  geom_density()
ggplot(data=differences, aes(x=diff_tens))+
  geom_density()
ggplot(data=differences, aes(x=diff_hundreds))+
  geom_density()
ggplot(data=differences, aes(x=diff_thousands))+
  geom_density()




```




```{r}
# Posteriors #, this is not complete, I copied and pasted from my blue sucker data.
# could be useful eventually when you want to simulate posterior predictions

posteriors <- as_draws_df(prior_test) %>% as_tibble() %>% expand_grid(length_s = length_s) %>%
  left_join(d %>% distinct(length_s, weight_s)) %>% 
  mutate(mu = exp(b_Intercept + b_weight_s*0 + b_length_s*length_s + 
                    `b_length_s:weight_s`*length_s*0),
         mu_prior = exp(prior_Intercept + prior_b_weight_s*0 + prior_b_length_s*length_s + 
                          `prior_b_length_s:weight_s`*length_s*0))

ggplot(posteriors_bsr_brm %>% filter(.draw <=1000), aes(x=length_s, y=mu_prior)) + 
  geom_point() +
  geom_line(aes(group  = .draw), alpha = 0.1)
```













